Supplementary Information

### 2020-08-26 doc/manuscript/MEE_manuscript.Rmd
added some content on cliamte -DBH interactions
#42

**Title:** Using tree-ring records to simultaneously characterize the influence of tree size, climate, and other environmental drivers on annual growth

@@

(*See [ISSUE #42 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/42)*)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
*(some details)* (Table S4).

**Table / figure on climate - DBH interactions?** (*See [ISSUE #42 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/42)*)

@@

Ideas to discuss:

* temperature sensitivity shifts from neg in warm climates to positive in cold climates, although @sniderhan_growth_2016 shows that the effect shifted to negative as warming progressed

* additive effects are common and should not be overlooked

@@

**We show that DBH is a strong driver of growth (any metric) relative to climate and that climate sensitivity often varies with DBH-- findings that have important implications for understanding changes in forest productivity in response to climate change.**
The observed importance of $DBH$ as a driver of tree growth reinforces the concept that changes in forest structure/ demography are critical to changes in forest productivity [REFS; @mcdowell_pervasive_2020].
The prevalence of interactive $DBH$-climate effects reinforces the concepts that drought tends to have a disproportionate impact on large trees [@bennett_larger_2015; @mcgregor_tree_nodate] and that *temperature sensitivity can vary across size classes (Rollinson et al. in review)*, implying that size should be considered in tree-ring analyses of climate sensitivity.
However, the relatively strong importance of $DBH$ in no way negates the importance of climate, as its effect would dominate changes in productivity in stands with relatively constant size structure.
Some studies have made a start at combining tree rings and forest census data to get at the climate sensitivity of whole-forest productivity [e.g., @helcoski_growing_2019], but have been limited in that they were not underlain by models of individual tree growth that simultaneously characterized the effects of tree size and climate, as we do here.
Further work will be required to combine this model with forest census data in order to understand the climate sensitivity of whole-forest productivity.

>>>>

**Title:** Using tree-ring records to simultaneously characterize the influence of tree size and climate drivers on annual growth

@@

*Interactive effects of climate and DBH*

We also ran GLS models testing for interactive effects of climate and DBH. For these, we modeled first-order linear relationships between $DBH$ and climate variables.
(*Need to give a better description here. See [ISSUE #42 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/42)*)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
Interactions were significant for over half of $DBH$-precipitation variable interactions for all three growth metrics (*51% for all*; Table S4).
The majority (#%) of these interactions were positive, indicating that larger trees generally respond more positively (or less negatively) to precipitation or it's frequency (Fig. **6?**).
Interactions between temperature variables and $DBH$ were significant for *~40%* of cases considered (Table S4). 
The majority (#%) of these interactions were positive, indicating that larger trees generally respond less negatively (or more positively) to $T_{max}$ or $PET$ (Fig. **6?**).

**figure on climate - DBH interactions?** (*See [ISSUE #42 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/42)*)

@@

Ideas to discuss:
* $T_{max}$ and $PET$ tend to be better variables than the more commonly used $T_{mean}$

* trees tend to be sensitive to water over longer time scales (makes sense-- lags caused by soil moisture storage)

* temperature sensitivity shifts from neg in warm climates to positive in cold climates (although @sniderhan_growth_2016 shows that the effect shifted to negative as warming progressed)

* additive effects are prevalent and should not be overlooked

@@

**We show that DBH is a strong driver of growth (any metric) relative to climate and that climate sensitivity often varies with DBH-- findings that have important implications for understanding changes in forest productivity in response to climate change.**
The observed importance of $DBH$ as a driver of tree growth reinforces the concept that changes in forest structure/ demography are critical to changes in forest productivity [REFS; @mcdowell_pervasive_2020].
The prevalence of interactive $DBH$-climate effects reinforces the concepts that large trees tend to be disproportionately sensitive to drought  [@bennett_larger_2015; @mcgregor_tree_nodate] but less sensitive to high temperatures *(Rollinson et al. in review)*, implying that size should be considered in tree-ring analyses of climate sensitivity.
However, the relatively strong importance of $DBH$ in no way negates the importance of climate, as its effect would dominate changes in productivity in stands with relatively constant size structure.
Some studies have made a start at combining tree rings and forest census data to get at the climate sensitivity of whole-forest productivity [e.g., @helcoski_growing_2019], but have been limited in that they were not underlain by models of individual tree growth that simultaneously characterized the effects of tree size and climate, as we do here.
Further work will be required to combine this model with forest census data in order to understand the climate sensitivity of whole-forest productivity.
